%%% The main file. It contains definitions of basic parameters and includes all other parts.

%% Settings for single-side (simplex) printing
% Margins: left 40mm, right 25mm, top and bottom 25mm
% (but beware, LaTeX adds 1in implicitly)
\documentclass[12pt,a4paper]{report}
\setlength\textwidth{145mm}
\setlength\textheight{247mm}
\setlength\oddsidemargin{15mm}
\setlength\evensidemargin{15mm}
\setlength\topmargin{0mm}
\setlength\headsep{0mm}
\setlength\headheight{0mm}
\setlength{\columnseprule}{0.2pt}
% \openright makes the following text appear on a right-hand page
\let\openright=\clearpage

%% Settings for two-sided (duplex) printing
% \documentclass[12pt,a4paper,twoside,openright]{report}
% \setlength\textwidth{145mm}
% \setlength\textheight{247mm}
% \setlength\oddsidemargin{14.2mm}
% \setlength\evensidemargin{0mm}
% \setlength\topmargin{0mm}
% \setlength\headsep{0mm}
% \setlength\headheight{0mm}
% \let\openright=\cleardoublepage

%% Generate PDF/A-2u
\usepackage[a-2u]{pdfx}

%% Character encoding: usually latin2, cp1250 or utf8:
\usepackage[utf8]{inputenc}
\usepackage{amssymb}
%% Prefer Latin Modern fonts
\usepackage{lmodern}

%% Further useful packages (included in most LaTeX distributions)
\usepackage{amsmath}        % extensions for typesetting of math
\usepackage{amsfonts}       % math fonts
\usepackage{amsthm}         % theorems, definitions, etc.
\usepackage{bbding}         % various symbols (squares, asterisks, scissors, ...)
\usepackage{bm}             % boldface symbols (\bm)
\usepackage{graphicx}       % embedding of pictures
\usepackage{fancyvrb}       % improved verbatim environment
\usepackage{natbib}         % citation style AUTHOR (YEAR), or AUTHOR [NUMBER]
\usepackage[nottoc]{tocbibind} % makes sure that bibliography and the lists
			    % of figures/tables are included in the table
			    % of contents
\usepackage{dcolumn}        % improved alignment of table columns
\usepackage{booktabs}       % improved horizontal lines in tables
\usepackage{paralist}       % improved enumerate and itemize
\usepackage{xcolor}         % typesetting in color
\usepackage{caption}			% for text alignment
\usepackage{listings}		% for code 

%%% Basic information on the thesis

% Thesis title in English (exactly as in the formal assignment)
\def\ThesisTitle{Artificial Intelligence for the Card Game Durak}

% Author of the thesis
\def\ThesisAuthor{Azamat Zarlykov}

% Year when the thesis is submitted
\def\YearSubmitted{2023}

% Name of the department or institute, where the work was officially assigned
% (according to the Organizational Structure of MFF UK in English,
% or a full name of a department outside MFF)
\def\Department{Department of Software and Computer Science Education (KSVI)}

% Is it a department (katedra), or an institute (ústav)?
\def\DeptType{Department}

% Thesis supervisor: name, surname and titles
\def\Supervisor{Adam Dingle, M.Sc.}

% Supervisor's department (again according to Organizational structure of MFF)
\def\SupervisorsDepartment{KSVI}

% Study programme and specialization
\def\StudyProgramme{Computer Science}
\def\StudyBranch{Artificial Intelligence}

% An optional dedication: you can thank whomever you wish (your supervisor,
% consultant, a person who lent the software, etc.)
\def\Dedication{%
I am filled with a sense of gratitude and admiration for my supervisor Adam Dingle for his unwavering support and guidance throughout this process. His constant patience and dedication to helping in difficult moments have been a constant source of inspiration. His expertise and knowledge along with his willingness to go above and beyond to provide assistance have provided me with valuable insights, especially during the challenging times of writing my thesis. I am truly fortunate and lucky to have had the opportunity to learn from a such remarkable mentor. 

Moreover, I would like to thank my family and friends who have always been there for me during the process of writing my thesis. Their encouragement, advice, and participation in playing the Durak game to analyze the strategies have been invaluable to me and I am so grateful for a constant source of strength, and wisdom and for simply being there to provide a shoulder to lean on when I needed it the most.
}

% Abstract (recommended length around 80-200 words; this is not a copy of your thesis assignment!)
\def\Abstract{%
Card games with imperfect information present a unique challenge for many common game-playing algorithms because of their hidden game state. The objective of this thesis is to create a framework for implementing and testing various AI agents in the popular imperfect information card game ``Durak'' to identify the most effective approach in this environment. This paper presents a theoretical and experimental comparison of agents using various techniques, including rules-based heuristics, minimax search, and Monte Carlo tree search. In our analysis, we found that the Monte Carlo Tree Search agent performed the best among the implemented AI agents, whereas the rule-based heuristic agent and the minimax agent were less effective.
}

% 3 to 5 keywords (recommended), each enclosed in curly braces
\def\Keywords{%
{artificial intelligence} {card game} {Durak}
}

%% The hyperref package for clickable links in PDF and also for storing
%% metadata to PDF (including the table of contents).
%% Most settings are pre-set by the pdfx package.
\hypersetup{unicode}
\hypersetup{breaklinks=true}

% Definitions of macros (see description inside)
\include{macros}

% Title page and various mandatory informational pages
\begin{document}
\include{title}

%%% A page with automatically generated table of contents of the bachelor thesis

\tableofcontents

%%% Each chapter is kept in a separate file
\include{preface}

\include{chapters/chap01}
\include{chapters/chap02}
\include{chapters/chap03}
\include{chapters/chap04}
\include{chapters/chap05}

\include{epilog}

%%% Bibliography
\include{bibliography}

%%% Figures used in the thesis (consider if this is needed)
\listoffigures

%%% Tables used in the thesis (consider if this is needed)
%%% In mathematical theses, it could be better to move the list of tables to the beginning of the thesis.
\listoftables

%%% Abbreviations used in the thesis, if any, including their explanation
%%% In mathematical theses, it could be better to move the list of abbreviations to the beginning of the thesis.
\chapwithtoc{List of Abbreviations}

%%% Attachments to the bachelor thesis, if any. Each attachment must be
%%% referred to at least once from the text of the thesis. Attachments
%%% are numbered.
%%%
%%% The printed version should preferably contain attachments, which can be
%%% read (additional tables and charts, supplementary text, examples of
%%% program output, etc.). The electronic version is more suited for attachments
%%% which will likely be used in an electronic form rather than read (program
%%% source code, data files, interactive charts, etc.). Electronic attachments
%%% should be uploaded to SIS and optionally also included in the thesis on a~CD/DVD.
%%% Allowed file formats are specified in provision of the rector no. 72/2017.
\appendix
\chapter{Attachments}

\section{User Documentation}
\label{userdoc}

The purpose of this attachment is to provide a user-friendly introduction to the game and its features. It is written in a way that is accessible to individuals without a technical background, enabling them to test the game without necessarily needing to develop AI for it. This attachment aims to furnish these users with the necessary knowledge and understanding of the game and its functions.

\subsection{Requirement}

In order to run various commands in Durak, the Windows 10 or Linux operating system is a necessary requirement. Additionally, the following software must also be installed: 

\section{Developer Documentation}
\label{devdoc}

\openright
\end{document}
